A Cross-Domain Benchmark for Active Learning

### 2024-06-04 doc/al_benchmark_paper.tex
Clarified semi-supervised setting

Here we step in with \emph{ALBench}, an active learning benchmark which covers
multiple application domains and reports a large amount of runs per
experiment, so that the significance of performance differences can can be
estimated. Specifically, \emph{ALBench} consists of datasets from
computer vision, natural language processing and the tabular domain. We provide
our datasets both in raw format %(i.e. consisting of text, images etc) 
as well as ``embedded'' by a fixed embedding model, enabling evaluation of AL methods in
this semi-supervised setting.
Furthermore, we propose two novel synthetic datasets to highlight general
challenges for AL methods.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{ALBench: A Cross-Domain Active Learning Benchmark}\label{sec:datasets}
A detailed description of the preprocessing of each dataset can be found in Appendix \ref{app:hyperparameters}. \\ [1mm]
\textbf{Tabular:}
AL research conducted on tabular data is sparse (only \cite{ashdeep} from the considered baseline papers). 
We, therefore, introduce a set of tabular datasets that we selected according to the following criteria:
(i) They should be solvable by medium-sized models in under 1000 samples, (ii) the gap between most AL algorithms and random sampling should be significant (potential for AL is present) and (iii) the gap between the AL algorithms and our oracle should also be significant (research on these datasets can produce further lifts).
We use \textbf{Splice}, \textbf{DNA} and \textbf{USPS} from LibSVMTools \cite{libsvmtools}.\\
\textbf{Image:}
We use \textbf{FashionMNIST} \cite{xiao2017fashion} and \textbf{Cifar10} \cite{krizhevsky2009learning}, since both are widely used in AL literature.\\
\textbf{Text:}
\todo{Schwache Erklärung}
We use \textbf{News Category} \cite{misra2022news} and \textbf{TopV2} \cite{chen-etal-2020-low-resource}.
Text datasets have seen less attention in AL research, but most of the papers that evaluate on text (\cite{hu2021towards}, \cite{zhou2021towards}) use at least one of these datasets.\\ [1mm]
%
We would like to point out that these datasets are selected for speed of computation (both in terms of the required classifier and the necessary budget to solve the dataset). 
However, we are solely focused on comparing different AL algorithms in this paper and do not aim to develop novel classification models on these datasets.
Our assumption is that a well-performing algorithm in our benchmark will also generalize well to larger real-world datasets, because we included multiple different data domains and classifier types and sizes in our experiments. \\ [1mm]
\todo{Muss die Verbindung zwischen encodings und semi-supervised learning klarer?}
Adapting the semi-supervised setting from \cite{hacohen2022active}, we offer all our datasets un-encoded (normal) as well as pre-encoded by a fixed embedding model that was trained by unsupervised contrastive learning. 
The text datasets are an exception to this, as they are only offered in their encoded form.
Pre-encoded datasets enable us to test small query sizes on more complex datasets like Cifar10 and FashionMnist.
They also serve the purpose of investigating the interplay between self-supervised learning techniques and AL, as well as alleviating the cold-start problem described in \cite{luth2024navigating} as they require a way smaller seed set.
The classification model for every encoded dataset is a single linear layer with softmax activation.
The embedding model was trained with the SimCLR \cite{chen2020simple} algorithm adopting the protocol from \cite{hacohen2022active}. 
%For Cifar10 and FashionMnist we adapt the reported hyperparameters from \cite{hacohen2022active} and for the tabular datasets we use random search to optimize the hyperparameters.
To ensure that enough information from the data is encoded by our embedding model, the quality of embeddings during pretext training was measured after each epoch.
To this end, we attached a linear classification head to the encoder, fine-tuned it to the data and evaluated this classifier for test accuracy. 
The checkpoint of each encoder model will be provided together with the framework. \\ [1mm]
Every dataset has a fixed size for the seed set of 1 sample per class, with the only exceptions being un-encoded FashionMnist and Cifar10 with 100 examples per class to alleviate the cold-start problem in these complex domains.

>>>>

Here we step in with \emph{ALBench}, an active learning benchmark which covers
multiple application domains and reports a large amount of runs per
experiment, so that the significance of performance differences can can be
estimated. Specifically, \emph{ALBench} consists of datasets from
computer vision, natural language processing and the tabular domain. We provide
our datasets both in normal format %(i.e. consisting of text, images etc)
as well as ``embedded'' by a fixed embedding model, enabling evaluation of AL methods in
this semi-supervised setting.
Furthermore, we propose two novel synthetic datasets to highlight general
challenges for AL methods.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{ALBench: A Cross-Domain Active Learning Benchmark}\label{sec:datasets}
A detailed description of the preprocessing of each dataset can be found in Appendix \ref{app:hyperparameters}. \\ [1mm]
\textbf{Tabular:}
AL research conducted on tabular data is sparse (only \cite{ashdeep} from the considered baseline papers). 
We, therefore, introduce a set of tabular datasets that we selected according to the following criteria:
(i) They should be solvable by medium-sized models in under 1000 samples, (ii) the gap between most AL algorithms and random sampling should be significant (potential for AL is present) and (iii) the gap between the AL algorithms and our oracle should also be significant (research on these datasets can produce further lifts).
We use \textbf{Splice}, \textbf{DNA} and \textbf{USPS} from LibSVMTools \cite{libsvmtools}.\\
\textbf{Image:}
We use \textbf{FashionMNIST} \cite{xiao2017fashion} and \textbf{Cifar10} \cite{krizhevsky2009learning}, since both are widely used in AL literature.\\
\textbf{Text:}
\todo{Schwache Erklärung}
We use \textbf{News Category} \cite{misra2022news} and \textbf{TopV2} \cite{chen-etal-2020-low-resource}.
Text datasets have seen less attention in AL research, but most of the papers that evaluate on text (\cite{hu2021towards}, \cite{zhou2021towards}) use at least one of these datasets.\\ [1mm]
%
We would like to point out that these datasets are selected for speed of computation (both in terms of the required classifier and the necessary budget to solve the dataset). 
However, we are solely focused on comparing different AL algorithms in this paper and do not aim to develop novel classification models on these datasets.
Our assumption is that a well-performing algorithm in our benchmark will also generalize well to larger real-world datasets, because we included multiple different data domains and classifier types and sizes in our experiments. \\ [1mm]
\todo{Muss die Verbindung zwischen encodings und semi-supervised learning klarer?}
Adapting the semi-supervised setting from \cite{hacohen2022active}, we offer all our datasets un-encoded (normal) as well as pre-encoded by a fixed embedding model that was trained by unsupervised contrastive learning (semi-supervised).
The text datasets are an exception to this, as they are only offered in their encoded form.
Pre-encoded datasets enable us to test small query sizes on more complex datasets like Cifar10 and FashionMnist.
They also serve the purpose of investigating the interplay between semi-supervised learning techniques and AL, as well as alleviating the cold-start problem described in \cite{luth2024navigating} as they require a way smaller seed set.
The classification model for every encoded dataset is a single linear layer with softmax activation.
The embedding model was trained with the SimCLR \cite{chen2020simple} algorithm adopting the protocol from \cite{hacohen2022active}. 
%For Cifar10 and FashionMnist we adapt the reported hyperparameters from \cite{hacohen2022active} and for the tabular datasets we use random search to optimize the hyperparameters.
To ensure that enough information from the data is encoded by our embedding model, the quality of embeddings during pretext training was measured after each epoch.
To this end, we attached a linear classification head to the encoder, fine-tuned it to the data and evaluated this classifier for test accuracy. 
The checkpoint of each encoder model will be provided together with the framework. \\ [1mm]
Every dataset has a fixed size for the seed set of 1 sample per class, with the only exceptions being un-encoded FashionMnist and Cifar10 with 100 examples per class to alleviate the cold-start problem in these complex domains.
